---
title: "Digital Communication, Part 2"
author: "Grace Rogers, Kelsey Sheehan, Taylor Ledford"
date: 'Updated: `r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, echo=FALSE}
library(tidyverse)
graces_bomb_theme <-  theme(panel.background = element_rect(fill = "grey70"))+
  theme(text=element_text(family = "Palatino"))
```

<hr>

1) In Part 1, you analyzed error rates for a single-channel system and a TRC system. One way to reduce transmission errors in either system would be to improve the initial 1% error rate. How much would reducing the error rate improve the performance of the TRC system? Create a plot that shows error rates from 0% to 1% (in increments of 0.1) on the x-axis versus the corresponding probability that any single TRC block (e.g., 000) is interpreted correctly (y-axis). To what value do we have to reduce the initial error rate in order for the probability of a block being correctly transmitted to be at least 99.99%?

```{r}
a <- seq(0, .01, by = .001)
ErrorsPerSec <- 10^9 * a

ggplot(NULL, aes(x = a, y = ErrorsPerSec)) +
  geom_bar(stat = "identity", color = "#F0A8A3", fill = "#F0A8A3") +
  theme_classic() +
  labs(title = "Performance of TRC by Error Rate", x = "Error Rate", y = "Error Per Second")

10^9 * .0001
print("The inital error needs to be reduced to .0001")
```

ANSWER: 


2) A 1000 Mbit/s system transmits 10^9^ bits per second. Each bit has a 0.03% chance of being incorrectly received when we use TRC and all bits are independent. There are are 86,400 seconds in a day. If X is the number of incorrect bits in a single second, use `rbinom` to simulate the distribution of X for a single day. Assess your simulation results for normality using a density plot, ECDF plot, QQ plot, and goodness of fit test. What do you conclude?

Simulation:
```{r}
prob <- rbinom(86400,10^9,.0003) %>%
  as.data.frame()

M <- mean(prob$.)
S <- sd(prob$.)

prob <- prob %>% mutate(normtestcats = case_when(. <= M - 2*S ~ "C1",
                                                 . <= M - 1*S & . > M - 2*S ~ "C2",
                                                 . <= M - 0*S & . > M - 1*S ~ "C3",
                                                 . <= M + 1*S & . > M - 0*S ~ "C4",
                                                 . <= M + 2*S & . > M + 1*S ~ "C5",
                                                 . >  M + 2*S ~ "C6"))
x <- prob$.
x.q <- quantile(prob$.)

ggplot(prob, aes(x = .)) + 
  geom_density() +
  stat_function(fun = dnorm, args = list(M, S), color = "#F5CB9B") +
  theme_classic() +
  labs(title = "Incorrect Bits for a Day", y = "Density", x = "Incorrect Bits")

ggplot(NULL, aes(x = x)) + 
  geom_step(stat = "ecdf") +
  stat_function(fun = pnorm, args = list(M, S), color = "#E0CDED") +
  theme_classic() +
  labs(title = "Incorrect Bits for a Day", y = "Fn(X)", x = "Incorrect Bits") +
  geom_vline(aes(xintercept=x.q),linetype = "dashed")

ggplot(prob, aes(sample = .)) + 
  stat_qq() +
  stat_qq_line(color = "#BACCE6") +
  theme_classic() +
  labs(title = "Incorrect Bits for a Day", x = "Theoretical", y = "Sample")
```

Ho: X is normal distributed.

Ha: X is not normal distributed.

Goodness of Fit Test:
```{r}
observed <- table(prob$normtestcats)

probs <- c(pnorm(-2), 
           pnorm(-1) - pnorm(-2), 
           pnorm(0) - pnorm(-1), 
           pnorm(1) - pnorm(0), 
           pnorm(2) - pnorm(1), 
           pnorm(2, lower.tail=F))

expected <- probs * sum(observed)

testStat <- sum((observed - expected)^2 / expected)
DF <- length(observed) - 3
pvalue <- pchisq(testStat, DF, lower.tail = FALSE)

print(testStat)
print(pvalue)
```

Conclusions: From the graphs X looks to be a strong normal distribution, and our p-value indicates that with a value greater than .1. 


<hr>

END PART 2