---
title: "Hypothesis Testing: Comparing Two Groups"
author: "Taylor Ledford"
date: "`r Sys.time()`"
output: html_document
---

Consider the Flight Delays dataset discussed in Chapter 1 of the Chihara & Hesterberg textbook, which contains data on selected flights departing from LaGuardia airport in 2009. Complete all problems below using that dataset. 

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
Delays <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/FlightDelays.csv")
```


### Problem 1

Conduct a traditional t-test and a permutation resampling test to determine whether the average delay times of the two airlines differ. 

```{r}
summary <- Delays %>%
  group_by(Carrier) %>%
  summarize(xbar = mean(Delay))

obsMeanDiff <- -diff(summary$xbar)

print(obsMeanDiff)

t.test(Delays$Delay ~ Delays$Carrier)

N <- 10^5 - 1
permMeanDiff <- numeric(N)

testdata <- Delays$Delay

for (i in 1:N){
   index <- sample(4029, 2906)
   AA <- testdata[index]
   UA <- testdata[-index]
   permMeanDiff[i] <- mean(AA) - mean(UA)
}

# ggplot(NULL, aes(x = permMeanDiff)) + 
#   geom_histogram(bins = 30, fill = "lightblue") + 
#   geom_vline(xintercept = obsMeanDiff, color = "red") +
#   labs(title = "Resampling Mean Differences")

# lower tail test with Ha: mu_b < mu_g

pvalue.lower <- (sum(permMeanDiff <= obsMeanDiff) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

pvalue.upper <- (sum(permMeanDiff >= obsMeanDiff) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

data.frame(alternative = c("Two-Tailed"),
           pvalue      = c(pvalue.twotail))
```



### Problem 2

Conduct a traditional z-test and a permutation resampling test to determine whether United Airlines has a greater proportion of flights delayed by 30 minutes than American Airlines. 

```{r}
summary2 <- Delays %>%
  group_by(Carrier) %>%
  summarize(n=n(), delayed=mean(Delayed30=="Yes"), notDelayed=mean(Delayed30=="No"))

print(summary2)

obsPropDiff <- -diff(summary2$delayed)
obsPropDiff

sprintf("UA has a delayed proportion greater than AA by %s", summary2[2, "delayed"] - summary2[1, "delayed"])

table2 <- table(Delays$Carrier, Delays$Delayed30)

prop.test(table2)

N <- 10^5 - 1
permPropDiff <- numeric(N)

testdata2 <- Delays$Delayed30

for (i in 1:N){
   index <- sample(4029, 2906)
   UA <- testdata2[index]
   AA  <- testdata2[-index]
   permPropDiff[i] <- mean(UA == "Yes") - mean(AA == "Yes")
}

ggplot(NULL, aes(x = permPropDiff)) + 
  geom_histogram(bins = 30, fill = "lightblue") + 
  geom_vline(xintercept = obsPropDiff, color = "red") +
  labs(title = "Resampling Proportion Differences")

# lower tail test with Ha: mu_b < mu_g

pvalue.lower <- (sum(permPropDiff <= obsPropDiff) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

pvalue.upper <- (sum(permPropDiff >= obsPropDiff) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

data.frame(alternative = c("Lower-Tail"),
           pvalue      = c(pvalue.lower))
```


### Problem 3

Conduct a traditional F-test and permutation resampling test to determine whether the variance in flight delay times is different between the two carriers. Unlike means and proportions, the test statistic used for a test of two variances is the _ratio_ of the variances, not the difference.

```{r}
summary3 <- Delays %>%
  group_by(Carrier) %>%
  summarize(sd=sd(Delay), variance=sd^2)

obsVarRatio <- summary3[1, "variance"] / summary3[2, "variance"]
obsVarRatio

table3 <- table(summary3$Carrier, summary3$variance)

var.test(Delay ~ Carrier, data = Delays)

N <- 10^5 - 1
permVarDiff <- numeric(N)

testdata3 <- Delays$Delay

for (i in 1:N){
   index <- sample(4029, 2906)
   UA <- testdata3[index]
   AA  <- testdata3[-index]
   permVarDiff[i] <- var(UA) / var(AA)
}

ggplot(NULL, aes(x = permVarDiff)) + 
  geom_histogram(bins = 30, fill = "lightblue") + 
  geom_vline(xintercept = obsVarRatio$variance, color = "red") +
  labs(title = "Resampling Population Variance")

# lower tail test with Ha: mu_b < mu_g

pvalue.lower <- (sum(permVarDiff <= obsVarRatio$variance) + 1) / (N + 1)

# upper tail test with Ha: mu_b > mu_g

pvalue.upper <- (sum(permVarDiff >= obsVarRatio$variance) + 1) / (N + 1)

# two-tail test with Ha: mu_b =/= mu_g

pvalue.twotail <- 2 * min(pvalue.lower, pvalue.upper)

data.frame(alternative = c("Two-Tailed"),
           pvalue      = c(pvalue.twotail))


```


<hr>

END!!!

