---
title: "Creating and Manipulating Vectors"
author: "Taylor Ledford"
date: "`r Sys.time()`"
output: html_document
---

### Problem 1

Create vectors to represent the outcomes or sample spaces for the following scenarios using `c`, `:`, `seq`, and/or `rep`. Whenever possible, try to be efficient by using the options other than `c`. Store your vectors using `<-` and use `print()` to show the contents.

* The sample space of a random variable X is {0, 1, 2, ..., 10}.

```{r}
library(plyr)
a <- 0:10
print(a)
```

* Scores on a test can be 0 or any positive, even integer up to 100.

```{r}
scores <- seq(from=0, to=100, by=2)
print(scores)
```

* A machine reads the pH of a solution to one decimal place of precision.

```{r}
pH <- seq(from=0, to=14, by=0.1)
print(pH)
```

* A player in a particular game can choose rock, paper, scissors, lizard, or Spock.

```{r}
RPSLZ <- c("rock", "paper", "scissors", "lizard", "Spock")
print(RPSLZ)
```

* A clock counts down from 60 seconds to zero seconds in whole-second increments.

```{r}
time <- seq(from=60, to=0)
print(time)
```

* A song's chorus contains the following three phrases, sung once in order then repeated.
    + I get knocked down
    + But I get up again
    + You're never gonna keep me down

```{r}
lyrics <- c("I get knocked down","But I get up again","You're never gonna keep me down")
rep(lyrics, times=2)
```

* In a particular group of 100 people, each person has one of four blood types: A, B, AB, or O.
    + A = 40%
    + B = 11%
    + AB = 4%
    + O = 45%
    
```{r}
BloodTypes <- c("A", "B","AB","O")
BloodPercent <- c("40%", "11%", "4%", "45%")
names(BloodPercent) <- BloodTypes
print(BloodPercent)
```


### Problem 2

Consider two classes of STT3850. Section 101 has 7 actuarial science majors, 12 computer science majors, 5 mathematics majors, and 7 majors in other science disciplines. The counts for Section 102 are 4, 13, 5, and 9. Be sure to `print()` each object you create below.

* Create two named vectors, one for each section, that contain the counts in each major with the appropriate major labels.
* Use the two vectors you made to create a third vector that contains the total number of students in each major category.
* Using the combined data, create a vector that shows the fraction of students in each major.
* Extract a vector that contains only the fractions for Actuarial Science and Mathematics.
* Create a data frame that contains the following variables: section, major, and count. 

```{r}
S101 <- c(7,12,5,7)
S102 <- c(4,13,5,9)
TotalSections <- append(S101, S102)
Sections <- c(101,101,101,101,102,102,102,102)
Majors <- c("Actuarial Science", "Computer Science", "Mathematics", "Other Sciences")
names(S101) <- Majors
names(S102) <- Majors
TotalMajors <- S101 + S102
MajorsPercent <- TotalMajors / sum(TotalMajors)
MathAndAct <- c(MajorsPercent["Actuarial Science"], MajorsPercent["Mathematics"])
MajorsDF <- data.frame("Section"=Sections, "Major"=Majors, "Count"=TotalSections)
print(MajorsDF)
```


### Problem 3

Suppose we want to simulate repeatedly rolling a single six-sided die.

* Create and `print()` vector called DIE that contains the integers 1 through 6.
* Use `sample()` to simulate rolling the die 1000 times and store the results as ROLLS. 
* Create and `print()` a summary to show how many times each value occurs in ROLLS.

```{r}
DIE <- c(1,2,3,4,5,6)
print(DIE)
ROLLS <- sample(x=DIE, size=1000, replace=TRUE)
ROLLS <- table(ROLLS)
print(ROLLS)
```


### Problem 4

Write a `for` loop to help create a vector that contains the first 20 numbers in the [Fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). 

* Use the `numeric` function to pre-create a vector of zeroes called FIB in which to store your results. 
* Set the first element of FIB to zero.
* Set the second element of FIB to one.
* Use a loop to calculate the remaining 18 elements (i.e., elements 3 through 20) and store them in FIB.
* Print the contents of FIB.
* If you don't get the correct 20 numbers, recheck your work!

```{r}
FIB <- numeric(length=20)
FIB[1] <- 0
FIB[2] <- 1
for (i in 3:20){
  FIB[i] <- FIB[i-1] + FIB[i-2]
}
print(FIB)
```

EXTENSION: Can you modify your code to allow someone to select how many numbers they want to generate?

```{r}
#n <- inputNumber
#FIB1 <- numeric(length=n)
#FIB1[1] <- 0
#FIB1[2] <- 1
#for (i in 3:n){
#  FIB1[i] <- FIB1[i-1] + FIB1[i-2]
#}
#print(FIB1)
```


<hr>

END!!!
